---
title: "Software Usage"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## CRAN package download stats

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(scales)
library(googlesheets4)
library(yaml)
library(cranlogs)
```

```{r, echo=FALSE}
download_stats <- cranlogs::cran_downloads(packages = c("ottrpal",
                                                        "conrad",
                                                        "ari",
                                                        "text2speech",
                                                        "metricminer"), #add more here within the parentheses
                                                        from = "2017-08-31", #first version of ari published on 2017-08-31
                                                        to = "last-day")
```

```{r, eval = FALSE, echo=FALSE}
download_stats_contributing <- cranlogs::cran_downloads(packages = c("spatialTIME"),
                                                        from = "2025-02-30", #date contribution was added to CRAN
                                                        to = "last-day")

download_stats <- rbind(download_stats, download_stats_contributing)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
download_stats_processed <- download_stats %>%
  separate(date, into=c("year", "month name", "day"), sep = "-") %>%
  unite("Month", c("year", "month name"), sep='-', remove=TRUE) %>%  
  group_by(Month, package) %>%
  summarise(monthly_downloads = sum(count)) %>% #summarize monthly downloads by package
  filter(monthly_downloads > 0) #drop the 0's
```

<!--
... plotting: Note that the color order is alphabetical

Alphabetically the packages are
- ari (color is `#440154`)
- conrad (color is `#2c728e`)
- metricminer (color is `#fde725`)
- ottrpal (color is `#28ae80`)
- spatialTime (when added, we'll use `#472d7b`)
- text2speech (color is `#addc30`)

Colors are viridis and have been grabbing options from here:
c("#fde725", "#addc30", "#5ec962", "#28ae80", "#21918c", "#2c728e", "#3b528b", "#472d7b", "#440154")
-->

### Over time

```{r echo = FALSE, fig.width=15.84, fig.height=6.39}

package_colors <- c("#440154", "#2c728e", "#fde725", "#28ae80" ,"#addc30" )
#package_colors <- c("#440154", "#2c728e", "#fde725", "#28ae80", "#472d7b" ,"#addc30" ) #adds in spatialTime

months_vec <- seq(as.Date("2017-08-01"), #set this reference date to the first day of hte month to make this vector work (even though ari was added at the end of the month)
                  as.Date(today()), by="months")

download_stats_processed %>%
  ggplot(aes(Month, monthly_downloads, group=package, color = package)) +
     geom_line() +
     geom_point() +
     scale_colour_manual(values=package_colors) +
     theme(panel.background = element_blank(),
           panel.grid = element_blank(),
           text = element_text(size = 17, family = "Arial")) +
     geom_vline(aes(xintercept = "2019-05"), linetype='dashed', linewidth = 1.5, color = '#addc30') + #text2speech published date
     geom_vline(aes(xintercept="2022-02"), linetype='dashed', linewidth = 1.5, color = '#28ae80') + #ottrpal published date
     geom_vline(aes(xintercept="2023-07"), linetype='dashed', linewidth = 1.5, color = '#2c728e') + #conrad published date
     geom_vline(aes(xintercept="2024-02"), linetype="dashed", linewidth = 1.5, color = '#fde725') + #metricminer published date
     #geom_vline(aes(xintercept="2025-01"), linetype="dashed", linewidth = 1.5, color="#472d7b") + #spatialTime published witwh contributions date
     theme(axis.text.x = element_text(angle = 90),
           legend.position = "bottom") +
     scale_x_discrete(breaks = format(months_vec[seq(1, length(months_vec), 2)], "%Y-%m")) + #every other month
     labs(x = NULL,
          y = "Monthly Downloads",
          color = "R Packages")
```

### Total downloads by package {.tabset}

#### For all time

```{r, echo= FALSE}
 DT::datatable(
      download_stats_processed %>%
      group_by(package) %>%
      summarize(total_downloads = comma(sum(monthly_downloads))),
      colnames = c("Package", "Total Downloads"),
      options = list(lengthChange = FALSE, # remove "Show X entries"
                     scrollY = "275px"), #may need to increase if you add packages and don't want to scroll
      # For the table to grow/shrink
      fillContainer = TRUE,
      escape = FALSE
    )
```

#### Past 6 months

```{r, echo=FALSE}
DT::datatable(
      download_stats_processed %>%
        filter(Month > format(as.Date(today() - months(6)), "%Y-%m")) %>% #within 6 months
        group_by(package) %>%
        summarize(total_downloads = comma(sum(monthly_downloads))),
      colnames = c("Package", "Total Downloads (last 6 months)"),
      options = list(lengthChange = FALSE, # remove "Show X entries"
                     scrollY = "275px"), #may need to increase if you add packages and don't want to scroll
      # For the table to grow/shrink
      fillContainer = TRUE,
      escape = FALSE
    )
```

### Total downloads overall

**Total downloads**, summed across packages, for all time:

```{r echo=FALSE, results = 'asis'}
cat(comma(unlist(download_stats %>% dplyr::summarize(download_total = sum(count)), use.names = FALSE)))
```


## GitHub Template Repositories

```{r echo = FALSE}
get_sync_repos <- function(read_yml_link){

  sync_file_contents <- read_yaml(read_yml_link)

  sync_repos <- sync_file_contents$group[[1]]$repos %>%
    str_trim() %>%
    as.data.frame() %>%
    separate_longer_delim(cols = ".", delim="\n") %>%
    separate(".", c("org", "repo"), sep = "/", remove = FALSE)

  return(sync_repos)
}
```

```{r echo=FALSE, results='asis'}
echo_num <- function(sync_df, template_name){
  cat(paste0("There are ", nrow(sync_df), " GitHub **epositories receiving updates from the ", template_name, "**. ", nrow(sync_df %>% filter(org == "jhudsl" | org == "fhdsl" | org == "ottrproject")), " of those are within the jhudsl, fhdsl, or ottrproject organizations"))
}
```

### OTTR Template - Courses

<!--
Main OTTR template
-->

```{r echo=FALSE, results='asis'}
sync_repos <- get_sync_repos("https://raw.githubusercontent.com/ottrproject/OTTR_Template/refs/heads/main/.github/sync.yml")
echo_num(sync_repos, "OTTR_Template")
```

### OTTR Template - Websites

<!--
OTTR website template
-->

```{r echo=FALSE, results='asis'}
web_sync_repos <- get_sync_repos("https://raw.githubusercontent.com/ottrproject/OTTR_Template_Website/refs/heads/main/.github/sync.yml")
echo_num(web_sync_repos, "OTTR_Template_Website")
```

<!--
### Metricminer Dashboard

```{r echo = FALSE, eval = FALSE, results='asis'}
dashboard_sync_repos <- get_sync_repos("https://raw.githubusercontent.com/ottrproject/metricminer-dashboard/refs/heads/main/.github/sync.yml")
echo_num(dashboard_sync_repos, "metricminer-dashboard)
```
-->

## Shiny Apps

### Loqui

<!--
Add the logo?
-->

```{r echo=FALSE, message=FALSE, warning-FALSE, results='hide'}
loqui_data <- sheet_results$loqui_data
loqui_data_gaps <- sheet_results$loqui_data_supp
```

<!--
Now (as of May 6 2025) fixed for data to this spreadsheet (previously wasn't as of July 2024)...Have timestamps (or placeholders) for videos that were produced with Loqui between those days/for which we don't have user data
-->

<!--
`loqui_data_gaps` contains a mixture of the word "Placeholder" (videos that were made with Loqui, but without any stored metadata/users didn't receive an email of the video) and with a date/time (videos that were made with Loquid, without any stored metadata besides a timestamp note in the logs because users were sent an email of the video)

`loqui_data` has a timestamp, a user email, and a video duration
-->

**Number of unique Loqui users**:

```{r echo=FALSE, results = 'asis'}
cat(length(unique(loqui_data$email)))
```

**Number of videos made with Loqui**:

```{r echo=FALSE, results = 'asis'}
cat(comma(nrow(loqui_data) + nrow(loqui_data_gaps)))
```

<!--
Want a graph of when loqui was used

The main dataset has a more "readable"/word format for dates, made with the blastula package's add_readable_time function (https://github.com/FredHutch/loqui/blob/160c3e820ccf2463c62375124c668be03a9e387f/app.R#L461, https://github.com/rstudio/blastula/blob/master/R/add_readable_time.R). Need to undo that action to combine it with the gaps/log data and plot
-->

```{r echo = FALSE, results = 'asis', message = FALSE, warning = FALSE}
loqui_data_dates <- loqui_data %>%
  separate(date_time, c("date", "time"), sep=" at ") %>%
  separate(date, c("weekday", "month_day", "year"), sep = ", ") %>%
  separate(month_day, c("month", "day"), sep = " ") %>%
  mutate(month = match(month, month.name)) %>%
  mutate(month = formatC(month, width = 2, format = "d", flag = "0"),
         day = formatC(as.integer(day), width = 2, format = "d", flag = "0")) %>%
  unite(date_time, c(year, month, day), sep = "-") %>%
  mutate(date_time = lubridate::ymd(date_time)) %>%
  select(date_time)

ldg <- loqui_data_gaps %>% filter(date_time != "Placeholder")
loqui_gaps_dates <- as.Date(lubridate::as_datetime(purrr::simplify(ldg$date_time))) %>% as_tibble() %>% `colnames<-`(c("date_time"))

binded_data <- bind_rows(
  loqui_data_dates,
  loqui_gaps_dates
)

count_df <- binded_data %>%
  group_by(date_time) %>%
  summarize(count = n())

alldays_df <- data.frame(date_time = ymd(seq.Date(as.Date(min(count_df$date_time)), as.Date(max(count_df$date_time)), by = "days")), count = 0)
daily_usage <- right_join(count_df, alldays_df, by="date_time") %>%
  mutate(daily_count = coalesce(count.x, count.y)) %>% #this is setting days without a timestamp to 0 and adding a 0 to days with responses
  select(!c(count.x, count.y)) %>%
  arrange(date_time)
toPlot <- daily_usage %>%
  separate(date_time, into=c("year", "month name", "day"), sep = "-") %>%
  unite("Month", c("year", "month name"), sep='-', remove=TRUE) %>%
  group_by(Month) %>%
  summarize(monthly_count = sum(daily_count))

highest_day_index <- which(daily_usage$daily_count == max(daily_usage$daily_count))  
```
<!--
Make a plot showing when loqui has been used? Does it get used a lot before semesters? Before the ITCR annual meeting? etc.
-->

```{r echo=FALSE, message = FALSE, warning = FALSE}
toPlot %>% ggplot(aes(Month, monthly_count, group = 1)) +
   geom_line() +
   geom_point() +
   theme(panel.background = element_blank(),
         panel.grid = element_blank(),
         text = element_text(size = 17, family = "Arial")) +
   #geom_vline(aes(xintercept = "2024-09"), linetype='dashed', linewidth = 1.5) + #ITCR Annual Meeting 2024
   theme(axis.text.x = element_text(angle = 90)) +
   #scale_x_discrete(breaks = format(months_vec[seq(1, length(months_vec), 2)], "%Y-%m")) + #every other month
   labs(x = NULL,
        y = "Monthly Count")
```

The highest single day usage of Loqui was on `r daily_usage[highest_day_index, "date_time"]`, with a total of `r daily_usage[highest_day_index, "daily_count"]` videos produced.

There have been a total of `r length(which(daily_usage$daily_count >= 25))` days where at least 25 videos were made with Loqui. 
